Inferring community characteristics in labelled networks

### 2021-05-19 nips/appendices.tex
All graphs loaded in

>>>>

\subsection{Algorithms}

\begin{algorithm} % enter the algorithm environment
	\caption{Block membership sample generation} % give the algorithm a caption
	\label{alg:b-samples} % and a label for \ref{} commands later in the document
	\begin{algorithmic} % enter the algorithmic environment

		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
		\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
		\IF{$\log \eta < \log \alpha_b$}
		\STATE $b^{(t+1)} \gets b'$
		\ELSE
		\STATE $b^{(t+1)} \gets b^{(t)}$
		\ENDIF
		\ENDFOR

		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$

	\end{algorithmic}
\end{algorithm}

\begin{algorithm} % enter the algorithm environment
	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
	\begin{algorithmic} % enter the algorithmic environment

		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$

		\item[]

		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
		\STATE $\xi \gets \sim \Gaussian(0, I)$
		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
		\IF{$\log \eta < \log \alpha_\theta$}
		\STATE $\theta^{(t+1)} \gets \theta'$
		\ELSE
		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
		\ENDIF
		\ENDFOR

		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$

	\end{algorithmic}
\end{algorithm}

### 2021-05-19 nips/experiments.tex
All graphs loaded in

\begin{itemize}
	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) - network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France. Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day. Vertex features include class membership, gender and whether or not the student is a teacher or pupil. These data were collected on consecutive days in October 2009. We choose to analyse just the second day. 
	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list. Each vertex has been manually labelled with a variety of features describing their relationship to the author. For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.
	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch. Vertex labels are extracted according to video-games played, location and streaming habits. This dataset is also broken down into disjoint networks according to language. We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S. politics, published close to the presidential election in 2004. Two books are connected if they were frequently co-purchased on Amazon. Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).
\end{itemize}

@@

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\fbox{\includegraphics[width=0.5\linewidth]{school-legend.png}}
		\caption{Graph coloured by inferred block memberships of each node}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-null.png}
		\caption{Feature-to-block generator sampled weight parameters for each block index}
		\label{fig:school-null}
	\end{subfigure}
	\caption{Primary school dynamic contacts network \cite{schools}}
\end{figure}

>>>>

\begin{itemize}
	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S. politics, published close to the presidential election in 2004. Two books are connected if they were frequently co-purchased by the same customer. Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).

	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) - network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France. Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day. Vertex features include class membership, gender and whether or not the student is a teacher or pupil. These data were collected on consecutive days in October 2009. We choose to analyse just the second day.

	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) - an assortment of Facebook egonets. These are networks of a particular user's friends list and all the connections within that. Vertex features are extracted from each user's profile and is fully anonymized. We focus on the egonet with id 1912.

	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list. Each vertex has been manually labelled with a variety of features describing their relationship to the author. For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.

	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.

	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch. Vertex labels are extracted according to video-games played, location and streaming habits. This dataset is also broken down into disjoint networks according to language. We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).

\end{itemize}

@@

\subsection{Political books}

This dataset was collected by \citet{polbooks}. We wish to determine whether political affiliation is a good predictor of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral). The inferred block memberships are given in figure \ref{fig:books-graph}.

\begin{figure}[!h]
	\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-graph.png}
		\fbox{\includegraphics[width=0.4\linewidth]{3-legend.png}}
		\caption{Graph coloured by inferred block memberships of each node}
		\label{fig:books-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-null.png}
		\caption{Feature-to-block generator sampled weight parameters for each block index}
		\label{fig:book-null}
	\end{subfigure}
	\caption{Political books network}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\fbox{\includegraphics[width=0.5\linewidth]{school-legend.png}}
		\caption{Graph coloured by inferred block memberships of each node}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-null.png}
		\caption{Feature-to-block generator sampled weight parameters for each block index}
		\label{fig:school-null}
	\end{subfigure}
	\caption{Primary school dynamic contacts network}
\end{figure}

\subsection{Facebook egonet}

\begin{figure}[!h]
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-graph.png}
		\fbox{\includegraphics[width=0.3\linewidth]{10-legend.png}}
		\caption{Graph coloured by inferred block memberships of each node}
		\label{fig:fb-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-null.png}
		\caption{Feature-to-block generator sampled weight parameters for each block index}
		\label{fig:fb-null}
	\end{subfigure}
	\caption{Primary school dynamic contacts network}
\end{figure}

### 2021-05-19 nips/inference.tex
intro
All graphs loaded in

So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample. This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance. This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain. The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$. We would rather deal with the expectation of each $y_{ij}^{(t)}$:
%
\begin{equation}
	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
	= p(b_i = j | A, X)
\end{equation}
%
We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
%
\begin{equation}
	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
\end{equation}
%
We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$. This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently. Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$. This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain. This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth. Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.

\begin{algorithm} % enter the algorithm environment
	\caption{Block membership sample generation} % give the algorithm a caption
	\label{alg:b-samples} % and a label for \ref{} commands later in the document
	\begin{algorithmic} % enter the algorithmic environment

		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
		\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
		\IF{$\log \eta < \log \alpha_b$}
		\STATE $b^{(t+1)} \gets b'$
		\ELSE
		\STATE $b^{(t+1)} \gets b^{(t)}$
		\ENDIF
		\ENDFOR

		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$

	\end{algorithmic}
\end{algorithm}

\begin{algorithm} % enter the algorithm environment
	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
	\begin{algorithmic} % enter the algorithmic environment

		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$

		\item[]

		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
		\STATE $\xi \gets \sim \Gaussian(0, I)$
		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
		\IF{$\log \eta < \log \alpha_\theta$}
		\STATE $\theta^{(t+1)} \gets \theta'$
		\ELSE
		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
		\ENDIF
		\ENDFOR

		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$

	\end{algorithmic}
\end{algorithm}

\subsection{Complexity Analysis}

>>>>

So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample. This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance. This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain. The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$. We would rather deal with the expectation of each $y_{ij}^{(t)}$:
%
\begin{equation}
	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
	= p(b_i = j | A, X)
\end{equation}
%
We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
%
\begin{equation}
	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
\end{equation}
%
We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$. This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently. Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$. This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain. This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; we refer to these as $T_b$ and $T_\theta$ henceforth. Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
